General Game Playing and Monte-Carlo Tree Search

### 2013-11-15 doc/paper/paper.tex
MORE

\documentclass[letterpaper]{article}
% Required Packages
\frenchspacing
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{listings}
\lstset{language=Lisp, basicstyle=\footnotesize}
\usepackage{caption}
\DeclareCaptionFont{white}{ \color{white} }
\captionsetup[lstlisting]{singlelinecheck=false, margin=0pt, font={bf,footnotesize}}
\usepackage{parcolumns}
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\nocopyright
% Title
\title{General Game Playing and Monte-Carlo Tree Search}
\author{Armando Ramirez \\ Purdue University, West Lafayette, IN \\ ramirez7@purdue.edu \\ armando.m.ramirez@gmail.com}
% Body
\begin{document}
\maketitle

@@

\subsubsection{Monte-Carlo Tree Search}
% TODO

@@

\subsubsection{Summary}
This paper presents extensions menat to be optimization for Monte-Carlo Tree Search that apply to any generally-described game.

\subsubsection{Critique}
% TODO

\subsection{LONG}

\subsubsection{Summary}
% TODO

\subsubsection{Critique}
% TODO

@@

\subsubsection{Summary}
% TODO

\subsubsection{Critique}
% TODO

\subsection{Synthesis}
% TODO
% Maybe I won't be able to write more than a few sentences
% In which case, remove the Synthesis section

@@

\begin{lstlisting}[frame=single,caption=TODO]
(<= (head & args)
    & tail)
;; -~->
(fresh [fresh-vars]
 (transformed head)
 (transformed tail))
\end{lstlisting}

\begin{lstlisting}[frame=single,caption=TODO]
(relation1)
(relation2)
;; ...
(relationN)
;; -~->
(conde
 [(transformed relation1)]
 [(transformed relation2)]
 ;; ...
 [(transformed relationN)])
\end{lstlisting}

@@

\subsubsection{Implementation}
My implementation is rather straightforward. It is modularized such that all the explicit steps discussed in a theoretical discussion of Monte-Carlo Tree Search are present and named in the implementation. To aid in programming this algorithm in Clojure, I used a simpler implementation for 3x3 Tic-Tac-Toe as reference \cite{randomcomp}. Because my implementation must work for any game described in GDL, there are some assumptions that can be made about 3x3 Tic-Tac-Toe that cannot be made in a General Game Playing MCTS agent:
\begin{itemize}
\item There can be arbitrary number of players. Therefore, the notion of win, loss, and draw are less relevant. Instead, the raw Goal relation score of each player is used.
\item There are not necessarily turns. If the game is synchronous, the search tree must expand according to all possible combinations of possible moves by all players.
\end{itemize}

\subsubsection{Results}
% TODO

>>>>

\documentclass[letterpaper]{article}
% Required Packages
\frenchspacing
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{listings}
\lstset{language=Lisp, basicstyle=\footnotesize}
\usepackage{caption}
\DeclareCaptionFont{white}{ \color{white} }
\captionsetup[lstlisting]{singlelinecheck=false, margin=0pt, font={bf,footnotesize}}
\usepackage{graphicx}
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\nocopyright
% Title
\title{General Game Playing and Monte-Carlo Tree Search}
\author{Armando Ramirez \\ Purdue University, West Lafayette, IN \\ ramirez7@purdue.edu \\ armando.m.ramirez@gmail.com}
% Body
\begin{document}
\maketitle

@@

\subsubsection{Monte-Carlo Tree Search}
Monte-Carlo Tree Search is a combination of the Monte-Carlo method of performing random simulations and game tree search. The algorithm traverses the game tree state by state, and performed random simulations at each state. As it traverses the game tree, the statistics are collected and combined. After the algorithm's budget is exhausted, these statistics are used to make an informed decision.

There are four canonical stages to Monte-Carlo Tree Search \cite{finnsson2012generalized}:
\begin{itemize}
\item \emph{Selection} The algorithm must choose a next-state amongst its explored children. A variety of selection algorithms are used to optimize this choice.
\item \emph{Playout} The Monte-Carlo simulation. A game is played to completion with each player making random moves.
\item \emph{Expansion} The tree begins with just the root node. Each simulation, the tree is expanded, usually by a single node.
\item \emph{Back-Propogation} After a simulation is performed or a terminal state is found, the results must be propogated to all the ancestors of the end state. This is because all its ancestors lead to the result.
\end{itemize}

These stages are visualized in the Figure 1 \cite{finnsson2012generalized}.

\begin{figure}[h]
\includegraphics[scale=0.65]{mctsfigure}
\centering
\caption{A single iteration of MCTS visualized}
\end{figure}

@@

\subsubsection{Summary}
This paper \cite{finnsson2012generalized} presents extensions menat to be optimization for Monte-Carlo Tree Search that apply to any generally-described game. It is important that this extensions to not make any assumptions about game structure or rules. These extensions aim to improve the execution time and accuracy of the search. Two of them affect the playout stage, which in many games can be quite costly. The final extension affects child selection, which is a key factor in how accurately the agent can traverse the tree.
\begin{itemize}
\item \emph{Goal Stability Early Cutoff} If a trend is confidently found in the progression of the Goal relation of a game, playout can be terminated early and the predicted winner can be stated as the winner.
\item \emph{Terminal Interval Early Cutoff} If a game has a tendency to end in an interval of turns, it is unneccessary to go more than 1/3 of the way through that interval \cite{finnbjorn}
\item \emph{Unexplored Action Urgency} ``Exploit the fringe of the MCTS tree'' by not always exploring all children in the selection phase if the already explored states give enough reason to move on from the current node.
\end{itemize}
% TODO more?

\subsubsection{Critique}

The extensions presented in the paper are all well-thought out and founded within the context of General Game Playing. An extremely conscious effort to avoid domain-specific assumptions is made, and in evaluating the extensions, the author is careful to choose a variety of games with a variety of qualities. The author presents actual results that show that across games, the extensions make nontrivial improvements to Monte-Carlo Tree Search.

However, in practice some of these extensions will end up accomplishing nothing. For example, in many games (such as Tic-Tac-Toe) there is a ternary Goal for every state: a player has either won, lost, or the game is in a state of draw. This goal will not fluctuate from draw until it reaches termination, so Goal Stability Early Cutoff is completely irrelevant to Tic-Tac-Toe.

For more complex games, the Goal Stability Early Cutoff extension is highly dependent on the encoding of the Goal state by the creator of the game description. For instance, in Chess, it would be reasonable to encode the Goal for a player as 100 if they have taken the enemy king, 0 if they have had their king taken, nad 50 if the game is drawn or still in progress. This goal would render the Goal Stability Early Cutoff extension powerless. That said, if the concept of material is introduced into the game, the extension would likely be very useful.

The Unexplored Action Urgency extension represents a key area of research and improvement for MCTS: Selection. In terms of both execution time and accuracy, more powerful Selection algorithms are key.

In the Related Work section, the author notes that some MCTS agents have used similar extensions before; however, \emph{none} have used such extensions without domain-specific knowledge. This paper is relatively recent (2012), and it is reasonable to expect research in General Game Playing Monte-Carlo Tree Search to continue to grow, especially with the recent introduction of GDL-II, which allows for stochastic and imperfect-information games \cite{StanfordGGP}.

\subsection{Understanding the Success of Perfect Information Monte Carlo Sampling in Game Tree Search}

\subsubsection{Summary}
This paper \cite{long2010understanding} is interesting in that it does not aim to explore improvements to MCTS. Instead, it aims to elucidate a previously unresolved problem: MCTS has a great deal of success playing imperfect information games, despite being fundamentally based in perfect information games.

The majority of the paper reasons about the generalized search tree itself, without regard for the actual steps of growing the tree. The paper outlines three basic properties of MCTS search trees for imperfect information games:
\begin{itemize}
\item \emph{Leaf Correlation} The probability that sibling nodes will have the same playoff value. The lower this is, the higher level of control a player has over their chances of winning the game no matter how late in the game it is.
\item \emph{Bias} Whether or not the game favors a player.
\item \emph{Disambiguation Factor} In imperfect information games, the disambiguation factor is how much the possible set of states shrinks as information is revealed.
\end{itemize}

The paper outlines three basic properties of MCTS search trees, and also explains ways to measure these properties at runtime. Once these properties can be measured, the paper goes on to explain how a MCTS agent that is aware of these properties can take advantage of them.

The author performed experiements using both generic synthetic game trees and real games. In both cases, the author's agent was able to take advantage of game tree properties to improve its payoff.

\subsubsection{Critique}

The author does excellent work formalizing properties of Monte-Carlo search trees as they apply to imperfect information games.

The synthetic trees used for testing represent two-player, zero-sum, stochastic imperfect information games. The assumptions of two-player and zero-sum are important and not intrinsic in the domain the paper is discussing, but such games are extremely common so it is natural for the author to use them for his synthetic trees. Despite these trees making these assumptions, they do allow the author to have fine-grained control over the three properties of Leaf Correlation, Bias, and Disambiguation Factor. This leads to a powerful model of MCTS that can be used for repeatable and understandable experiments.

Thankfully, the author did not purely work with synthetic game trees. Real games were used as well to test the ideas presented in the paper. In addition, in real games, the properties must be measured at runtime, whereas in synthetic trees they are known ahead of time. By presenting examples of real games, the author greatly substantiates that his claims operate outside the user-designed synthetic game tree realm.

The author did not discuss General Game Playing; however, with the advent of GDL-II for stochastic and imperfect information games, it is likely that the concepts core to this paper will be used in GDL-II competitions. GDL-II is fairly recent, so it is a vast, unexplored problem space, but in the coming years research such as this author's will be applied to it.

In the conclusion, the author points out that Perfect Information MCTS agents can likely be taken advantage of by actually being predictable in its mistakes. This topic itself merits further discussion, as an agent that actively takes advantage of Perfect Information MCTS would be hugely successful in the General Game Playing arena.

@@

\subsubsection{Summary}

\subsubsection{Critique}
This paper is the most specific of the three critiqued. It deals specifically in Go, and less specifically in two-player, perfect information, \emph{zero-sum} games. General Game Playing makes no assumptions about games being zero-sum, so this paper making that assumption greatly limits the domain in which its findings are applicable.

@@

\begin{lstlisting}[frame=single,caption={\textless= relations are joined in a fresh block}]
(<= (head & args)
    & tail)
;; -~->
(fresh [fresh-vars]
 (transformed head)
 (transformed tail))
\end{lstlisting}

\begin{lstlisting}[frame=single,caption=Multiple related relations are joined by disjunction]
(relation1)
(relation2)
;; ...
(relationN)
;; -~->
(conde
 [(transformed relation1)]
 [(transformed relation2)]
 ;; ...
 [(transformed relationN)])
\end{lstlisting}

@@

\subsubsection{Implementation}
My implementation is rather straightforward. It is modularized such that all the explicit steps discussed in a theoretical discussion of Monte-Carlo Tree Search are present and named in the implementation. To aid in programming this algorithm in Clojure, I used a simpler implementation for 3x3 Tic-Tac-Toe as reference \cite{randomcomp}. Because my implementation must work for any game described in GDL, there are some assumptions that can be made about 3x3 Tic-Tac-Toe that cannot be made in a General Game Playing MCTS agent:
\begin{itemize}
\item There can be arbitrary number of players. Therefore, the notion of win, loss, and draw are less relevant. Instead, the raw Goal relation score of each player is used.
\item There are not necessarily turns. If the game is synchronous, the search tree must expand according to all possible combinations of possible moves by all players.

\end{itemize}

The search is itself modularized into an \texttt{iteration} function that can be called a desired amount of times, accumulating statistics with each iteration. This allows for the user to be able to tightly control the amount of resources are used by the search.

\subsubsection{Results}
With a budget as low as 10 iterations, my Monte-Carlo Tree Search algorithm was able to consistently (80\%+ winrate) beat a random player at Tic-Tac-Toe. Its execution time was approximately 2 seconds per turn when running on ECE 207 Redhat computers.

For other games, the execution time was far too long to be reasonable (in large part due to the lengthy playout stages of the games.)
